---
title: "The analysis of COVID-19 in United States"
author: "J. Terao"
date: '2022-11-07'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

### Set URLs
```{r get_jhu_data, message=FALSE}
## Get current Data in the four files
# they all begin the same way
library(tidyverse)
library(lubridate)
url_in <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"
file_names <- c("time_series_covid19_confirmed_global.csv", "time_series_covid19_deaths_global.csv", "time_series_covid19_confirmed_US.csv", "time_series_covid19_deaths_US.csv")
urls <- str_c(url_in, file_names)
```

### Import Data
```{r import_data, message=FALSE}
global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])
US_cases <- read_csv(urls[3])
US_deaths <- read_csv(urls[4])
```
### Make Data Tidy
```{r tidy_global_data, message=FALSE}
global_cases <- global_cases %>%
    pivot_longer(cols = -c(`Province/State`, `Country/Region`, Lat, Long), names_to = "date", values_to = "cases") %>%
    select(-c(Lat, Long))

global_deaths <- global_deaths %>%
    pivot_longer(cols = -c(`Province/State`, `Country/Region`, Lat, Long), names_to = "date", values_to = "deaths") %>%
    select(-c(Lat, Long))
```

```{r tidy_US_data, message=FALSE}
US_cases <- US_cases %>%
    pivot_longer(cols = -(UID:Combined_Key), names_to = "date", values_to = "cases") %>% select(Admin2:cases) %>% mutate(date = mdy(date)) %>% select(-c(Lat, Long_))
US_deaths <- US_deaths %>%
    pivot_longer(cols = -(UID:Population), names_to = "date", values_to = "deaths") %>% select(Admin2:deaths) %>% mutate(date = mdy(date)) %>% select(-c(Lat, Long_))
```

```{r combine_global, message=FALSE}
global <- global_cases %>% full_join(global_deaths) %>% rename(Country_Region = `Country/Region`, Province_State = `Province/State`) %>% mutate(date = mdy(date))

global <- global %>% unite("Combined_Key", c(Province_State, Country_Region), sep = ", ", na.rm = TRUE, remove = FALSE)

US <- US_cases %>% full_join(US_deaths)
```

```{r uid_lookup, message=FALSE}
uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"
uid <- read_csv(uid_lookup_url) %>% select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))
```

```{r fill_global, message=FALSE}
global <- global %>% left_join(uid, by = c("Province_State", "Country_Region")) %>% select(-c(UID, FIPS)) %>% select(Province_State, Country_Region, date, cases, deaths, Population, Combined_Key)
```

## Visualizing Data
### Preparing
```{r us_by_state, message=FALSE}
US_by_state <- US %>% 
  group_by(Province_State, Country_Region, date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths), Population = Population)%>% 
  mutate(deaths_per_mill = deaths * 1000000 / Population) %>%
  select(Province_State, Country_Region, date, cases, deaths, deaths_per_mill, Population) %>%
  ungroup()
```

```{r US_totals, message=FALSE}
US_totals <- US_by_state %>%
  group_by(Country_Region, date) %>%
  summarize(cases = sum(cases), deaths = sum(deaths),
            Population = sum(Population)) %>%
  mutate(deaths_per_mill = deaths * 1000000 / Population) %>%
  select(Country_Region, date, cases, deaths, deaths_per_mill, Population) %>%
  ungroup()
```

```{r new_cases, message=FALSE}
US_by_state <- US_by_state %>% 
  mutate(new_cases = cases - lag(cases), 
         new_deaths = deaths - lag(deaths))
US_totals <- US_totals %>%
  mutate(new_cases = cases - lag(cases), 
         new_deaths = deaths - lag(deaths))

US_by_state <- na.omit(US_by_state)
US_totals <- na.omit(US_totals)

#tail(US_totals %>% select(new_cases, new_deaths, everything()))
```

### Visualization of the prevention rate of death
```{r find_best_and_worst, message=FALSE}
US_state_totals <- US_by_state %>%
  group_by(Province_State) %>% 
  summarize(deaths = max(deaths), cases = max(cases), 
            population = max(Population),
            cases_per_thou = 1000 * cases / population, 
            deaths_per_thou = 1000 * deaths / population, 
            deaths_per_cases = 1000 * deaths / cases) %>% 
  filter(cases > 0, population > 0)

US_state_totals %>% 
  ggplot(aes(x = cases_per_thou, y = deaths_per_thou, label = Province_State)) +
  geom_text() +
  labs(title = "Comparing the result of death prevention", y = "Deaths per population", x = "Cases per population")

```


### Visualization of the best and worst state of the prevention rate of death
```{r compare_best_and_worst, message=FALSE}
state <- "North Carolina"
US_by_state_min <- US_by_state %>% 
  filter(Province_State == state) %>% 
  filter(new_cases > 0) %>%
  filter(new_deaths >= 0)

plot <-
  ggplot(data = US_by_state_min, mapping = aes(x = date, y = new_cases)) + 
  geom_line(aes(color = "new_cases")) + 
  geom_point(aes(color = "new_cases")) +
  geom_line(aes(y = new_deaths, color = "new_deaths")) + 
  geom_point(aes(y = new_deaths, color = "new_deaths")) + 
  scale_y_log10() + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  labs(title = str_c("COVID19 in ", state), y = NULL)

plot

state <- "Mississippi"
US_by_state_max <- US_by_state %>% 
  filter(Province_State == state) %>% 
  filter(new_cases > 0) %>%
  filter(new_deaths >= 0)

plot <-
  ggplot(data = US_by_state_max, mapping = aes(x = date, y = new_cases)) + 
  geom_line(aes(color = "new_cases")) + 
  geom_point(aes(color = "new_cases")) +
  geom_line(aes(y = new_deaths, color = "new_deaths")) + 
  geom_point(aes(y = new_deaths, color = "new_deaths")) + 
  scale_y_log10() + 
  theme(legend.position = "bottom", 
        axis.text.x = element_text(angle = 90)) + 
  labs(title = str_c("COVID19 in ", state), y = NULL)

plot
```

## Model
 We have the hypothesis that there is correlation between cases per population and deaths per population. However, the state of North Carolina has lower deaths per population than other state. This result shows the government of North Carolina might prevent severe symptoms of COVID-19. Our hypothesis should be considered well.

## Bias
 Some state have less observation than others in the same period. It could make time-series analysis ambiguous. 